---
title: "Lesson 2"
author: "Gaetano Scaduto"
output: slidy_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
```

# Let's work with a real dataset

-   R is based on an open source philosophy.

-   Some things you have *included in the basic package*

-   You can download other items from a [online archive](https://cran.r-project.org/) called the C-RAN (The Comprehensive R Archive Network).

-   These are "**packages**" that allow you to expand the basic functionality of R.

---

Before using a package for the first time, **only for the first time**, you must install it.

-   To install it, you need an **internet connection** and to run the command ***install.packages()***.

-  We are now going to use three packages

-   **rio**, **ggplot2**, and **texreg**

To install packages, use the following syntax.

```{r, eval=FALSE}
install.packages("rio")
install.packages("ggplot2")
install.packages("texreg") #Note: this, after the hash sign, is a comment.
#Comments are portions of text that the computer does not look at; they are used by us humans to enrich the code 
#with explanations and notes.
```

---

After installing, you must call the library.

-   This must be done **every time you restart R** with the **library()** command.

```{r, warning = F}

library(rio)
library(ggplot2)
library(texreg)
```

One last refinement. I want numbers with decimal points, not scientific notation.

To do this, I tell R this.

```{r, warning = F}
options(scipen = 999)
```

# Importing a dataframe

The rio library (which stands for R Input/Output) allows you to import files into R without any hassle.

In our case, I have prepared a dataset with the **results of the 2022 elections in all Italian municipalities**, plus some other information.

-   This dataset is based on [open data from the Ministry of the Interior](https://elezionistorico.interno.gov.it/), which I manipulated a little and combined with some statistics from ISTAT.

---

To import it into R, you need to do the following:

-   Create a folder somewhere on your computer and give it a name.

-   Put the dataset uploaded online into that folder.

-   You can also just copy and paste it into that folder..

-   After that, you will need to tell R that you are working in that folder.

    -   To do this, you will need to execute the ***setwd()*** command (which stands for "set working directory") and enter the **folder path** in quotation marks.**.

        -   To find the folder path, if you have Windows, simply right-click on the folder and then select 'Copy as path'.

        -   If you have a Mac, search Google/AI for 'how to copy Mac folder path', but I think you can see it in 'properties' by right-clicking or something like that.

        -   Once you have set the working directory, all the files you save from R will be found there!

            -   If you really can't manage it: **exit R and** **double-click on the dataset file**. If you want to learn how to use R, this isn't a good idea, but just this once, it's fine...

---

-   **Please note! If you are using Windows, what will be copied will look like this:**

    -   "C:\\Users\\gasca\\OneDrive - Università della Strada\\Desktop\\cartella esempio"

    -   R per uno strano motivo vuole che le stanghette non siano "\\" ma "/", quindi voi dovrete cambiarle manualmente e scrivere invece

        ```{r, eval=FALSE}
        setwd("C:/Users/gaetanoscaduto/OneDrive - Università della Strada/Desktop/cartella esempio")
        ```

Quindi il comando sarà:

```{r, eval = FALSE}

setwd("C:/Users/gaetanoscaduto/OneDrive - Università della Strada/Desktop/cartella esempio")
```

```{rm echo=F}
setwd("C:/Users/100722gsc/OneDrive - Erasmus University Rotterdam/Postdoc/Teaching/Laboratorio-R/Lezione 2")
```

---

Once you have done this, run the following command (do not forget the quotation marks "") to import the dataframe into your environment.

```{r, warning = F}
data = import("ris2022_lezione.RDS")
```

-   NOTE: if you double-clicked on the dataset from the folder (for those who did not know how to copy the folder path), do not execute the above command!

In this case, you are telling R:

-   Take the object named in the way inside the quotation marks (from the folder we mentioned earlier), import it into R and call it, from now on, "data".

-   "data" is the name we will use to refer to our dataframe.

-   Please note that you must always include the name of the file you wish to import, **including the extension (.RDS)**.

-   With the ***import()*** function (which is part of rio), you can easily import virtually any type of existing data file (this include .dta, .csv, .sav, .xlsx, and so on).

---

Note: this dataset has some missing data. For simplicity, we will work with the complete data. I will therefore remove all rows that have "missing data" with this command.

```{r, warning = F}
data = data[complete.cases(data), ]
```

# Let's take a peek at the data

Now we want to take a peek at the data. We can do this with the command ***head().***

```{r, warning = F}
head(data)
```

If you want to see the data in more detail, run the command ***View(data)*** on R and you can freely navigate the dataset.

---

If we want a quick summary of what the dataset contains, we can run the following command.

```{r, warning = F}
summary(data)
```

---

Now let's take a look at some specific statistics.

-   To do this, we will need to directly "call" the columns of the dataset.

  -   The operator ***\$*** is used after the dataset name (e.g. ***data\$istat***).

    -   Rstudio will then suggest the names of the available columns so that you can access them.

Let's calculate some basic statistics as an example.

-   Like the average number of residents in the municipalities in our dataset.

    -   To calculate the mean, use the command ***mean()**.*

```{r, warning = F}
mean(data$num_residenti)
```

Or the median, with the command ***median()***

```{r, warning = F}
median(data$num_residenti)
```

---

Or we ask: "In the municipality where the Five Star Movement received the highest percentage, how much did it get?"

-   The answer is given by the ***max()*** command on the variable (column) that represents the percentage of votes for the Five Star Movement (in our case, the variable "M5S").

```{r, warning = F}
max(data$M5S)
```

What if we want the minimum instead? We use the ***min()*** command.

```{r}
min(data$M5S)
```

---

Now we wonder whether, for example, the average number of votes obtained by Forza Italia in the municipalities of Calabria differs from that in the municipalities of Umbria.

```{r, warning = F}
mean(data[data$regione == "Calabria", ]$Fi)

mean(data[data$regione == "Umbria", ]$Fi)
```

# Tables

-   Tables can be very useful for understanding how our data is distributed.

-   We can create tables with one variable or with two variables, and in both cases we use the "table" command.

  -   Let's see, for example, how the variable ***areeint*** is distributed.

    -   (Istat internal area code - indicates the remoteness of the municipality: A is a central municipality, F is a remote municipality).

```{r, warning = F}
table(data$areeint)
```

---

-   If we want to create a table of two crossed variables, just put them in the ***table()*** command separated by a comma.

```{r, warning = F}
table(data$macroarea, data$areeint)

```

This table is not very informative, so we can consider creating a table with proportions using the following command.

```{r, warning = F}
prop.table(table(data$macroarea, data$areeint))
```

---

It's impossible to understand! Too many decimal places!

-   We solve this with the "round" function, which allows us to round the decimals to as many digits as we want.

    -   We only want two digits.

```{r, warning = F}
round(prop.table(table(data$macroarea, data$areeint)), digits = 2)
```

---

-   These are the percentages across the entire table.

  -   Perhaps it would be more useful to have them by row or by column.

    -   To do this, we need to add an option to prop.table.

-   1 corresponds to the percentages per row.

-   2 to the percentages per column.

```{r, warning = F}
round(prop.table(table(data$macroarea, data$areeint), 1), digits = 2)

```

```{r, warning = F}
round(prop.table(table(data$macroarea, data$areeint), 2), digits = 2)

```

# Regression!

-   Finiamo la nostra carrellata con qualche regressione.

-   Sappiamo tutti cos'è la regressione?

    -   For those who need a refresher, I made a quick guide for another course a few years ago (in Italian). You can download it here. [qui](https://github.com/gaetanoscaduto/teaching/blob/main/Regressione%20PP/Dispensa.pdf).

  -   Here we will only perform linear regressions.

-   Let's suppose we want to try to build a model that explains the vote for the Five Star Movement.

---

-   Linear regressions are performed using the "lm" command.

    -   The first *argument* of "lm" is ***data=*** followed by the name of your dataset (which in our case is also called ***data***).

    -   A comma must then be placed to separate the first and second *arguments.*

    -   Then put the dependent variable, your y

    -   After that, insert a tilde \~ (Alt+126).

    -   Finally, the independent variables.

    -   Save the entire result in an object

    -   Observe the results of your regression by placing the object in a ***screenreg()*** function.

-   NOTE: ***screenreg()*** only works if you have called the *texreg* library first with the ***library()*** command!

---

-   Let's start with a very famous model: "the Five Star Movement are the party of the South".

```{r, warning = F}
regr1 = lm(data=data, M5S ~ macroarea)
screenreg(regr1)

#Notice: these are conditional means! You could also run:
#mean(data[data$macroarea == "South", ]$M5S) - mean(data[data$macroarea == "Center", ]$M5S) 
```


Quick question: what is the reference category?

```{r}
#Notice: these are conditional means! You could also run:
mean(data[data$macroarea == "South", ]$M5S) - mean(data[data$macroarea == "Center", ]$M5S) 
```

---

Now let's try a more sophisticated model, which also takes peripherality into account.

``` {r, warning = F}
regr2 = lm(data=data, M5S ~ macroarea + areeint)
screenreg(regr2)
```

---

Finally, let's also try to include the number of residents, which is a continuous variable, and the area of the municipality, which is also a continuous variable.

```{r, warning = F}
regr3 = lm(data=data, M5S ~ macroarea + areeint + num_residenti + superficie)
screenreg(regr3)
```

---

This concludes our second part. Once again, do not be alarmed. R is challenging. My advice is to be patient, do not get frustrated, and make the most of artificial intelligence to understand errors and how it works. After a while, you will succeed, and it will be worth it.
